---
title: "Data Types (solution)"
output:
  html_document:
    df_print: paged
  github_document: 
    df_print: tibble  
---

<!-- This file by Charlotte Wickham is licensed under a Creative Commons Attribution 4.0 International License, adapted from the orignal work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->

```{r setup}
library(tidyverse)
library(lubridate)

# Example of a factor
eyes <- factor(x = c("blue", "green", "green"), 
               levels = c("blue", "brown", "green"))

# An example for times/dates
library(fivethirtyeight)
births <- US_births_1994_2003 %>%
  select(date, births)
```

## Warm-up / Review

Using the data `gss_cat`, find the average hours of tv watched (`tvhours`) for each category of marital status (`marital`).

```{r}
gss_cat %>%
  group_by(marital) %>% 
  summarise(avg_tv = mean(tvhours))
```

## Your Turn 1

What kind of object is the `marital` variable?  

```{r}
gss_cat
```

The `marital` variable is a **factor**.  

Brainstorm with your neighbor, all the things you know about that kind of object.

* Factors are R's way of storing categorical data.
* They have a set of allowable values called **levels**.
* Internally they are stored as integers.

# Factors

## Your Turn 2

Fix your summary of average hours of tv watched (`tvhours`) by marital status (`marital`), to ignore missing values, 
then create a plot to examine the results.

```{r}
gss_cat %>%
  drop_na(tvhours) %>%
  group_by(marital) %>% 
  summarise(avg_tvhours = mean(tvhours)) %>%
  ggplot() +
    geom_point(aes(avg_tvhours, marital))
```

## Your Turn 3

Fill in the blanks (`___`) to explore the average hours of tv watched by religion.

```{r}
gss_cat %>%
  drop_na(tvhours) %>%
  group_by(relig) %>%
  summarise(avg_tvhours = mean(tvhours)) %>%
  ggplot() +
    geom_point(mapping = aes(x = avg_tvhours, 
      y = fct_reorder(relig, avg_tvhours)))
```

## Quiz 

Why is this plot not very useful?

```{r}
gss_cat %>%
  drop_na(tvhours) %>%
  group_by(denom) %>%
  summarise(avg_tvhours = mean(tvhours)) %>%
  ggplot() +
    geom_point(mapping = aes(x = avg_tvhours,
      y = fct_reorder(denom, avg_tvhours)))
```

There are too many categories, and the categories are poorly labelled.

## Your Turn 4

Edit the code to also relabel some other Baptist denominations:

* "Baptist-dk which"    
* "Other baptists"  

```{r}
gss_cat %>%
  mutate(denom = fct_recode(denom,
    "Baptist - Southern" = "Southern baptist",
    "Baptist - Don't know" = "Baptist-dk which",
    "Baptist - Other" = "Other baptists")
  ) %>%
  pull(denom) %>%
  levels()
```

## Your Turn 5

What does the function `detect_denom()` do?

```{r}
detect_denom <- function(x){
  case_when(
    str_detect(x, "[Bb]ap") ~ "Baptist", 
    str_detect(x, "[Pp]res") ~ "Presbyterian",
    str_detect(x, "[Ll]uth") ~ "Lutheran",
    str_detect(x, "[Mm]eth") ~ "Methodist",
    TRUE ~ x
  )
}

gss_cat %>% pull(denom) %>% levels() %>% detect_denom()
```

It tries to detect a higher level grouping by looking for patterns in the denomination strings.

This is ideal for combining with `fct_revel()`:
```{r}
gss_cat %>%
  drop_na(tvhours) %>%
  mutate(denom_higher = fct_relabel(denom, detect_denom) %>%
      fct_reorder(tvhours, mean)) %>%
  group_by(denom_higher, denom) %>%
  summarise(avg_tvhours = mean(tvhours)) %>%
  ggplot() +
    geom_point(mapping = aes(x = avg_tvhours,
      y = denom_higher)) 
```

# Strings

With your neighbor, predict what these might return:

```{r}
strings <- c("Apple", "Pineapple", "Orange")

str_detect(strings, pattern = "pp")
str_detect(strings, pattern =  "apple")
str_detect(strings, pattern = "[Aa]pple")
```

Then run them!

# Times and Dates

## Your Turn 7

For each of the following formats (of the same date), pick the right `ymd()` function to parse them:

```{r}
ymd("2018 Feb 02")
mdy("2-1-18")
dmy("01/02/2018")
```

## Your Turn 8

Fill in the blanks to:

* Extract the month from date. 
* Extract the year from date.
* Find the total births for each year/month.
* Plot the results as a line chart.

```{r}
births %>%
  mutate(year = year(date),
    month = month(date, label = TRUE)) %>%
  group_by(year, month) %>%
  summarise(total_births = sum(births)) %>%
  ggplot() + 
    geom_line(aes(x = month, y = total_births, group = year))
```

# Take Aways

Dplyr gives you three _general_ functions for manipulating data: `mutate()`, `summarise()`, and `group_by()`. Augment these with functions from the packages below, which focus on specific types of data.

Package   | Data Type
--------- | --------
forcats   | factors
stringr   | strings
lubridate | dates and times
hms       | times